---
title: "Assignment 3 Task 3"
author: "Jonah Seif"
date: "11/3/2020"
output: html_document
---

```{r setup, echo=TRUE, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(effsize)
```

## Read in sea urchins data and clean names

```{r, message = FALSE}
urchins_clean <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Exploratory data visualization comparing the distribution of sizes between purple and red urchins
- Faceted histogram 
- Boxplot with common name on the x-axis and size on y-axis

```{r}
ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram(binwidth = .5)+
  facet_wrap(~common_name)+
  labs(x = "Urchin Diameter (cm)", y = "Count")


ggplot(data = urchins_clean,
       aes(x = common_name, y = size))+
  geom_boxplot()+
  labs(x = "Common Name", y = "Urchin Diameter (cm)")
```

## Create a summary table of statistics for red and purple urchins
- Contains: mean, median, standard deviation, sample size

```{r, message=FALSE}
urchin_stats <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size),
            sd = sd(size),
            median = median(size),
            n = n())

urchin_stats

urchins_purp <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  pull(size)

urchins_red <- urchins_clean %>% 
  filter(common_name == "Red Urchin") %>% 
  pull(size)

cohen.d(urchins_purp, urchins_red)
```
Based on the table and graphs above, we see that purple urchins have a narrower size distribution and a lower mean size compared to red urchins. The histograms show a fairly normal distribution for red urchins and a narrower, close to normal distribution for purple urchins, but with a positive tail. The standard deviation numbers from our table also demonstrate that the red urchin sample population has a greater spread of sizes compared to the purple urchin sample population. Specifically, the absolute difference of the means is 2.31cm (41.1% difference) and the Cohen's d is 1.49, which would indicate a noticeable difference between mean urchin diameters. 


## Explore purple urchin counts by site and year
- Filter to only include observations for purple urchins
- Convert the date to class "Date"
- Create a new column that only contains the year
- Create a new column that contains the full name for each of the five sites

```{r}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reed"))
```


## Make a summary table for total number of purple urchins observed by site and year

```{r}
purple_count <- purple_urchins %>% 
  group_by(year, site_full) %>% 
  count()

purple_count
```

## Make an exploratory graph of purple urchin numbers by site and year

```{r}
purple_count %>% ggplot(aes(x = year, y = n, fill = site_full))+
  geom_col(aes(color = site_full))+
  scale_fill_manual(values = c("red", "yellow", "green", "cyan", "purple"))+
  facet_wrap(~site_full)+
  theme(legend.position = "none")+
  labs(x = "Year", y = "Count")
```












